Inefficient Concessions and Mediation

### 2020-12-29 notes.tex
add calculations for We > W1, correct P5 in notes

\vskip.3in
Here $W1 > W0$ and $g1 > g0$. Welfare under mediator is never positive. 
\begin{itemize}
	\item T = 7, W = 8, D = 5, p = .5, dl = .25, dh = .6
\end{itemize}

\vskip.3in
Here, $W0 > W1 > We$ (mediator). Mediator optimally chooses e=0 anyway
\begin{itemize}
	\item T = 7, W = 8, D = 5, p = .3, dl = .4, dh = .9
\end{itemize}

\vskip.3in
Here, $W0 > W1 > We$ (mediator). Mediator optimally chooses $e=0.03$
\begin{itemize}
	\item T = 6, W = 8, D = 5, p = .5, dl = .2, dh = .9
\end{itemize}

						\vskip.3in
						Welfare comparisons (always for high type)
							\begin{itemize}
								\item When $e=1$: $\frac{1}{1-d_h}\left\{pT(1+g^*) + (1-p)(W-D(1+g^*))\right\} - \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D}$
									\begin{itemize}
										\item \color{red} Parameterization in the paper: welfare is 7.18; gift is 0.63
										\item \color{orange} Parameterization 2: $g^1$ is negative, creates negative upper bound for $d_h$. Separating with concessions equilibrium can't be sustained.
									\end{itemize}
								\item When $e=0$: $\frac{1}{1-d_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$
									\begin{itemize}
										\item \color{red} Parameterization in the paper: welfare is 8; gift is 3
										\item \color{orange} Parameterization 2: welfare is 5.7; gift is 10.8
									\end{itemize}
								\item When $0<e<1$ under a mediator: $\frac{1}{1-d_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - g_M$
									\begin{itemize}
										\item Optimal $g_M$: Expression~\ref{eq:g2} creates a menu of $(e,g)$ pairs that satisfy the constraint:
											\[
												g_M \geq \frac{\left(1 - \de \right) \left( T + D \right)}{p - e \left( p + T - \de T + W \right)}
											\]
										\item We're assuming $p - e \left( p + T - \de T + W \right) > 0$ (assuming gifts must be non-negative)
										\item Also $pg - \left(1 - \de \right) \left( T + D \right) \geq 0$ ($e$ version; this is WEAKER than the condition on $e$)
										\item \color{red} For the parameterization in the paper, welfare is negative for $e=0$ and $e=0.01$, the only values of $e$ that lead to non-negative gifts; gifts range from 52.5 to 218.75	
										\item \color{orange} Parameterization 2: welfare is 10.083 at optimal gift-efficiency $e=0.06$, $g = 116.67$, $eg = 7$
											\begin{itemize}
												\item Compared to $e=0$, welfare (gross of gift) goes up from 16.5 to 126.75. The entire difference is due to the positive effect of the material value of the gift (the full term is $\frac{1}{1-\de_h}pTeg$). The effective gift is $7$, which is much larger than the 1 in $\frac{1}{1-\de_h}pT\cdot 1$ that is there in both terms.
												\item The increase in the size (cost) of the gift from 10.8 to 116.67 is smaller than the increase in gross-of-gift welfare
											\end{itemize}
									\end{itemize}
							\end{itemize}

@@

When $0<e<1$ under a mediator, the optimal $g_M$ is the one that chooses the highest welfare from a menu of $(e,g)$ pairs:
$$\frac{1}{1-d_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)}$$

When $e=0$, welfare is $$\frac{1}{1-d_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$

We compare them to find when the welfare under the mediator is larger:
	\begin{multline*}
		\frac{1}{1-d_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} \geq \\
		\frac{1}{1-d_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$
	\end{multline*}							
Eliminating common terms:
	\begin{equation*}
		\frac{1}{1-d_h}\left\{pTeg_M \right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} \geq - p(T+D)
	\end{equation*}	
Rearranging:
	\begin{equation*}
		\frac{pTeg_M}{1-d_h} \geq \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} - p(T+D)
	\end{equation*}	
Here we can clearly see the increase in welfare from the impact of the effective gift on the left has to be greater than the increase in the size of the gift.
	\begin{itemize}
		\item \color{blue} I'm not sure anything reasonable can come from this
	\end{itemize}

\subsubsection{To do and Notes}		
	\color{ForestGreen} 
			\begin{itemize}
					\item NEXT STEP: check condition for sustainable $e=1$ but prefer mediator with $e < 1$ (I've found parameterizations where $e=1$ is preferred so far, but I derived an expression in my notes for when mediator is better $$g^M(pTe - 1) \geq g^1(pT + pD - D - 1)$$
						\begin{itemize}
							\item I'll probably need to type these notes up
						\end{itemize}
					\item Can I derive an analytical expression for when the mediator enhances welfare? See Section~\ref{sec:WM0}
						\begin{itemize}
							\item This is likely (surely?) to be when efficient gifts can't be sustained as a CSE.
							\item If you can sustain $e=1$ with no mediator, would you ever want a mediator?
						\end{itemize}
				\item When $g^1$ turns negative, is that exactly when $e=1$ can't be sustained? If we can get positive $g^1$, not sure we'd need the mediator? It's not about $g^1$, but the $\de_h^1$ condition.
					\begin{itemize}
						\item When $g^1$ is negative, the constraint on $\de_h^1$ (see proof of Thm 3) becomes an upper bound. If it's negative, it means that no $\de_h$ exists to make separating possible.
						\item I need a formal result to see clearly what is going on here
					\end{itemize}
			\end{itemize}

>>>>

\vskip.2in
Parameterization 3: $W1 > W0$ and $g1 > g0$. Welfare under mediator is never positive. 
\begin{itemize}
	\item $T = 7, \ W = 8, \ D = 5, \ p = .5, \ \de_l = .25, \ \de_h = .6$
\end{itemize}

\vskip.2in
Parameterization 4: $W0 > W1 > We$ (mediator). Mediator optimally chooses e=0 anyway
\begin{itemize}
	\item $T = 7, \ W = 8, \ D = 5, \ p = .3, \ \de_l = .4, \ \de_h = .9$
\end{itemize}

\vskip.2in
Parameterization 5: $W1 > W0 > We$ (mediator). Mediator optimally chooses $e=0.03$
\begin{itemize}
	\item $T = 6, \ W = 8, \ D = 5, \ p = .5, \ \de_l = .2, \ \de_h = .9$
\end{itemize}

						\vskip.3in
						Welfare comparisons (always for high type)
							\begin{itemize}
								\item When $e=1$: $\frac{1}{1-\de_h}\left\{pT(1+g^*) + (1-p)(W-D(1+g^*))\right\} - \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D}$
									\begin{itemize}
										\item \color{red} Parameterization in the paper: welfare is 7.18; gift is 0.63
										\item \color{orange} Parameterization 2: $g^1$ is negative, creates negative upper bound for $\de_h$. Separating with concessions equilibrium can't be sustained.
									\end{itemize}
								\item When $e=0$: $\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$
									\begin{itemize}
										\item \color{red} Parameterization in the paper: welfare is 8; gift is 3
										\item \color{orange} Parameterization 2: welfare is 5.7; gift is 10.8
									\end{itemize}
								\item When $0<e<1$ under a mediator: $\frac{1}{1-\de_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - g_M$
									\begin{itemize}
										\item Optimal $g_M$: Expression~\ref{eq:g2} creates a menu of $(e,g)$ pairs that satisfy the constraint:
											\[
												g_M \geq \frac{\left(1 - \de \right) \left( T + D \right)}{p - e \left( p + T - \de T + W \right)}
											\]
										\item We're assuming $p - e \left( p + T - \de T + W \right) > 0$ (assuming gifts must be non-negative)
										\item Also $pg - \left(1 - \de \right) \left( T + D \right) \geq 0$ ($e$ version; this is WEAKER than the condition on $e$)
										\item \color{red} For the parameterization in the paper, welfare is negative for $e=0$ and $e=0.01$, the only values of $e$ that lead to non-negative gifts; gifts range from 52.5 to 218.75	
										\item \color{orange} Parameterization 2: welfare is 10.083 at optimal gift-efficiency $e=0.06$, $g = 116.67$, $eg = 7$
											\begin{itemize}
												\item Compared to $e=0$, welfare (gross of gift) goes up from 16.5 to 126.75. The entire difference is due to the positive effect of the material value of the gift (the full term is $\frac{1}{1-\de_h}pTeg$). The effective gift is $7$, which is much larger than the 1 in $\frac{1}{1-\de_h}pT\cdot 1$ that is there in both terms.
												\item The increase in the size (cost) of the gift from 10.8 to 116.67 is smaller than the increase in gross-of-gift welfare
											\end{itemize}
									\end{itemize}
							\end{itemize}

@@

When $0<e<1$ under a mediator, the optimal $g_M$ is the one that chooses the highest welfare from a menu of $(e,g)$ pairs:
$$\frac{1}{1-\de_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)}$$

When $e=0$, welfare is $$\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$

We compare them to find when the welfare under the mediator is larger:
	\begin{multline*}
		\frac{1}{1-\de_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} \geq \\
		\frac{1}{1-\de_h}\left\{pT + (1-p)(W-D)\right\} - p(T+D)$$
	\end{multline*}							
Eliminating common terms:
	\begin{equation*}
		\frac{1}{1-\de_h}\left\{pTeg_M \right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} \geq - p(T+D)
	\end{equation*}	
Rearranging:
	\begin{equation*}
		\frac{pTeg_M}{1-\de_h} \geq \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} - p(T+D)
	\end{equation*}	
Here we can clearly see the increase in welfare from the impact of the effective gift on the left has to be greater than the increase in the size of the gift.
	\begin{itemize}
		\item \color{blue} I'm not sure anything reasonable can come from this
	\end{itemize}

\subsubsection{Welfare under Mediator vs. e=1}
\label{sec:WM1}
Here I'll try to derive a general result for when the welfare under a mediator would be higher than under $e=1$ separating with concessions equilibrium. \\

When $0<e<1$ under a mediator, the optimal $g_M$ is the one that chooses the highest welfare from a menu of $(e,g)$ pairs:
$$\frac{1}{1-\de_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)}$$

When $e=1$, welfare is $\frac{1}{1-\de_h}\left\{pT(1+g^1) + (1-p)(W-D(1+g^1))\right\} - \frac{p(1-d_l)(D+T)}{(1-d_l)(1-pT)+(1-p)D}$ \\

We compare them to find when the welfare under the mediator is larger:
	\begin{multline*}
		\frac{1}{1-\de_h}\left\{pT(1+eg_M) + (1-p)(W-D)\right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} \geq \\
		\frac{1}{1-\de_h}\left\{pT(1+g^1) + (1-p)(W-D(1+g^1))\right\} - \frac{p(1-d_l)(T+D)}{(1-d_l)(1-pT)+(1-p)D}
	\end{multline*}

Eliminating common terms:
	\begin{multline*}
		\frac{1}{1-\de_h}\left\{pTeg_M \right\} - \frac{\left(1 - \de_l \right) \left( T + D \right)}{p - e \left( p + T - \de_l T + W \right)} \geq \\
		\frac{1}{1-\de_h}\left\{pTg^1 + (1-p)(- Dg^1))\right\} - \frac{p(1-d_l)(T+D)}{(1-d_l)(1-pT)+(1-p)D}
	\end{multline*}			
Rearranging:
	\begin{equation*}
		\frac{1}{1-\de_h}\left\{pTeg_M \right\} - g_M \geq \frac{1}{1-\de_h}\left\{pTg^1 + (1-p)(- Dg^1))\right\} - g^1
	\end{equation*}

\color{blue} I think it's possible that this condition can never be satisfied (i.e. if e=1 is IC, you'd never want to call in the mediator). But I haven't tried to prove it yet

\subsubsection{To do and Notes}		
	\color{ForestGreen} 
			\begin{itemize}
					\item NEXT STEP: check condition for sustainable $e=1$ but prefer mediator with $e < 1$ 
						\begin{itemize}
							\item I've found parameterizations where $e=1$ is preferred (P5)
							\item Try to find a parameterization that shows the reverse (counterexample to my hypothesis that if $e=1$ is sustainable, you never invoke mediator) or see if I can prove that it never happens
						\end{itemize}
					\item Can I derive an analytical expression for when the mediator enhances welfare? See Section~\ref{sec:WM0}
						\begin{itemize}
							\item This is likely (surely?) to be when efficient gifts can't be sustained as a CSE.
							\item If you can sustain $e=1$ with no mediator, would you ever want a mediator?
						\end{itemize}
				\item When $g^1$ turns negative, is that exactly when $e=1$ can't be sustained? If we can get positive $g^1$, not sure we'd need the mediator? It's not about $g^1$, but the $\de_h^1$ condition.
					\begin{itemize}
						\item When $g^1$ is negative, the constraint on $\de_h^1$ (see proof of Thm 3) becomes an upper bound. If it's negative, it means that no $\de_h$ exists to make separating possible.
						\item I need a formal result to see clearly what is going on here
					\end{itemize}
			\end{itemize}
